---
title: "grpSLOPE ADMM"
output:
  pdf_document: default
  html_notebook: default
---

For now ADMM for `grpSLOPE` is implemented in an alternate (experimental) branch in the github repository.

To install the `grpSLOPE` version from that branch on github do:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
devtools::install_github("agisga/grpSLOPE", ref = "optimization_and_ADMM")
```

Then load the package:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(grpSLOPE)
library(tictoc) # for measuring runtime of ADMM vs FISTA
```

Here is a basic demonstration of how ADMM can be used with grpSLOPE.

Generate some data, and define grouping structures and prior weights.

```{r}
set.seed(2018)
n <- 200
p <- 400
n_grp <- 100
# design matrix
A <- matrix(rnorm(n * p), n, p)
# group membership per predictor
grp <- sample(x = 1:n_grp, size = p, replace = TRUE)
# weights
wt_per_grp <- sqrt(c(table(grp))) # i.e. sqrt of the group size as weight for each group
wt_per_predictor <- wt_per_grp[as.character(grp)]
# a sparse vector of coefficients per predictor (i.e., the true solution)
grp_signif <- 1:10
ind_signif <- which(grp %in% grp_signif)
x <- rep(0, p)
x[ind_signif] <- rnorm(length(ind_signif), mean = 5, sd = 1)
# response variable
y <- A %*% x + rnorm(n)
# normalize
col_norms <- apply(A, 2, function(a) sqrt(sum((a - mean(a))^2)))
A <- scale(A, center = TRUE, scale = col_norms)
y <- scale(y, center = TRUE, scale = FALSE)
```

Apply ADMM to solve the Group SLOPE optimization problem:
_(the ADMM algorithm is not yet included in the user-friendly function `grpSLOPE()`;
instead, it has to be called explicitly, and with more input parameters)_

```{r}
# regularizing parameters
lambda <- lambdaGroupSLOPE(method = "corrected", fdr = 0.1, n.obs = n,
                           group = grp, wt = wt_per_grp)

# get Group SLOPE results with ADMM
tic()
result_admm <- admmSolverGroupSLOPE(y = y, A = A, group = grp,
                                    wt = wt_per_predictor, lambda = lambda,
                                    rho = 1, infeas.tol = 1e-6, verbose = FALSE)
toc()
```

Apply the proximal gradient method to solve the Group SLOPE optimization problem:

```{r}
tic()
result_fista <- proximalGradientSolverGroupSLOPE(y = y, A = A, group = grp,
                                                 wt = wt_per_predictor,
                                                 lambda = lambda, dual.gap.tol = 1e-6,
                                                 infeas.tol = 1e-6, verbose = FALSE)
toc()
```

*As we can see, in this case, ADMM is 2.5 times as fast as FISTA. However, the speed of ADMM largely depends on the choice of the parameter `rho`.*

Compare the solutions:

```{r}
summary(result_admm$x - result_fista$x)
```

As expected, they are the same up to about 1e-6.

Compare the obtained results to the true solution:

```{r}
print(paste("TP:", sum(x != 0 & result_admm$x != 0), ";", 
            "FP:", sum(x == 0 & result_admm$x != 0)))
```

Note that the above computations, strictly speaking, don't follow the full approach of [Brzyski et. al. 2017], because the orthogonalization within groups is not performed.

Using the full approach of Brzyski et. al. (with FISTA optimization), we get:

```{r}
tic()
result <- grpSLOPE(X = A, y = y, group = grp, fdr = 0.1, sigma = 1)
toc()

print(paste("TP:", sum(x != 0 & result$beta != 0), ";",
            "FP:", sum(x == 0 & result$beta != 0)))
```

